On the Limitations of Weighted Model Counting

### 2020-04-01 paper.tex
Add back the polyadic stuff and add some notes

Previous/related work:
\begin{itemize}
\item Hailperin's approach to probability logic
  \cite{DBLP:journals/ndjfl/Hailperin84}
\item Nilsson's (somewhat successful) probabilistic logic
  \cite{DBLP:journals/ai/Nilsson86}
\item Logical induction: a huge paper with a good overview of previous attempts
  to assign probabilities to logical sentences in a way that makes sense
  \cite{DBLP:journals/eccc/GarrabrantBCST16}.
\item Semiring programming \cite{DBLP:journals/corr/BelleR16}
\item WMI \cite{DBLP:conf/ijcai/BellePB15}
\item Measures on Boolean algebras: more specific articles
  \begin{itemize}
  \item On possibility and probability measures in finite Boolean algebras
    \cite{DBLP:journals/soco/CastineiraCT02}
  \item Representation of conditional probability measures
    \cite{krauss1968representation}
  \end{itemize}
\end{itemize}

@@

\begin{definition} \label{def:wmc}
  Let $\mathbf{B}$ be an atomic BA, and let $M \subset \mathbf{B}$ be its set of
  atoms. Let $L \subset \mathbf{B}$ be such that every atom $m \in M$ can be
  uniquely expressed as $m = \bigwedge L'$ for some $L' \subseteq L$, and let
  $w\colon L \to \mathbb{R}_{\ge 0}$ be arbitrary. The \emph{weighted model
    count} $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ is defined as
  \[
    \WMC_w(x) = \begin{cases}
      0 & \text{if } x = 0 \\
      \prod_{l \in L'} w(l) & \text{if } M \ni x = \bigwedge L' \\
      \sum_{\text{atoms } a \le x} \WMC_w(a) & \text{otherwise}
    \end{cases}
  \]
  for any $x \in \mathbf{B}$. Furthermore, we define the \emph{normalised
    weighted model count} $\nWMC_w\colon \mathbf{B} \to [0, 1]$ as $\nWMC_w(x) =
  \frac{\WMC_w(x)}{\WMC_w(1)}$ for all $x \in \mathbf{B}$. For both $\WMC_w$ and
  $\nWMC_w$, we will drop the subscript when doing so results in no potential
  confusion.
%  Finally, we say that a measure $m\colon \mathbf{B} \to
%  \mathbb{R}_{\ge 0}$ is a \emph{WMC measure} (or is \emph{WMC-computable}) if
%  there exists a subset $L \subset \mathbf{B}$ and a weight function $w\colon L
%  \to \mathbb{R}_{\ge 0}$ such that $m = \WMC_w$.
\end{definition}
% TODO: mention that the definition can be reduced to a single formula (i.e.,
% without cases)
% TODO: any measure is a WMC measure if all atoms are in L

@@

\begin{proposition}[Measures on quotients] \label{def:measures_on_quotients}
  Let $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ be a measure, and let $(p)$ be
  a principal ideal. Let $m^*\colon \mathbf{B}/(p) \to \mathbb{R}_{\ge 0}$ be
  defined as
  \[
    m^*(a/(p)) = m(a \lor p)
  \]
  for any $a \in \mathbf{B}$. The function $m^*$ is well-defined. Furthermore,
  it is a measure on $\mathbf{B}/(p)$ if and only if $m(p) = 0$. Moreover, if it
  is a measure, then the following properties transfer from $m$ to $m^*$:
  \begin{itemize}
  \item if $m$ is a \emph{probability measure}, then so is $m^*$;
  \item if $m$ is \emph{strictly positive}, then so is $m^*$.
  \end{itemize}
\end{proposition}
\begin{proof}
  \Cref{lemma:well-defined} proves that the function is well-defined. Next,
  note that
  \[
    m^*(0/(p)) = m(0 \lor p) = m(p),
  \]
  so $m^*(0/(p)) = 0$ if and only if $m(p) = 0$. For the second part of
  \cref{def:measure}, let $a/(p), b/(p) \in \mathbf{B}/(p)$ be such that
  \[
    a/(p) \land b/(p) = a \land b / (p) = 0 / (p).
  \]
  This condition is equivalent to $a \land b \in (p)$ and $(a \land b) \lor p =
  p$ by well-known properties of quotients and ideals
  \cite{givant2008introduction}, \cref{def:ideal}, and the definition of $\le$,
  respectively. Now
  \begin{align*}
    m^*(a/(p) \lor b/(p)) &= m^*(a \lor b / (p)) = m(a \lor b \lor p) = m((a \lor p) \lor (b \lor p)) \\
                          &= m(a \lor p) + m(b \lor p) - m((a \lor p) \land (b \lor p)) \\
                          &= m^*(a/(p)) + m^*(b/(p)) - m((a \lor p) \land (b \lor p))
  \end{align*}
  by \cref{lemma:sum}. However
  \[
    (a \lor p) \land (b \lor p) = (a \land b) \lor p = p,
  \]
  so $m^*(a/(p) \lor b/(p)) = m^*(a/(p)) + m^*(b/(p))$ if and only if $m(p) = 0$.

@@

@@

\begin{theorem}[Using the previous results]
  Let $L \subset \mathbf{B}$ and $w\colon L \to \mathbb{R}_{\ge 0}$ be such that
  $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge0}$ is a WMC measure. Let $p \in
  \mathbf{B}$ be such that $\WMC(p) = 0$, and let $g\colon \mathbf{B}/(p) \to
  \mathbf{B}$ be defined as $g(a/(p)) = a \land \neg p$. Then $\WMC \circ g$
  is a WMC measure on $\mathbf{B}/(p)$ with respect to the same $L$ and $w$
  such that
  \[
    \begin{tikzcd}
      \mathbf{B} \arrow{d}[swap]{\WMC} & \mathbf{B}/(p) \arrow{l}[swap]{g}
      \arrow{dl}{\WMC \circ g} \\
      \mathbb{R}_{\ge 0}
    \end{tikzcd}
  \]
  commutes.
\end{theorem}

% TODO: how can WMC compute something in the quotient algebra?
% TODO: prove that the two ways to compute the probability give the same answer,
% i.e., how B(Delta) = B(L)/(not Delta).
% TODO: need an example of how we can compute probabilities in the quotient
% algebra.
% TODO: we can then move through the quotient BA using representatives we get
% from mapping true things to 1 and false things to 0.

@@

@@

% TODO: a section on WMI: 3 results

>>>>

Potential directions to explore (no good ones, really):
\begin{itemize}
\item Infinite BAs.
  \begin{itemize}
  \item For example, the BA of finite and cofinite sets could be interesting.
  \item OUWMC requires infinite logics to be compact (whatever that means). My
    algebraic angle suggests that completeness should be enough. Topologically,
    compactness implies completeness, but I have no idea how this translates to
    logics and algebras.
  \end{itemize}
\item WMI
\item Measures with something other than $\mathbb{R}_{\ge 0}$ as the codomain
  (doesn't look promising).
\end{itemize}

Previous/related work:
\begin{itemize}
\item Hailperin's approach to probability logic
  \cite{DBLP:journals/ndjfl/Hailperin84}
\item Nilsson's (somewhat successful) probabilistic logic
  \cite{DBLP:journals/ai/Nilsson86}
\item Logical induction: a big paper with a good overview of previous attempts
  to assign probabilities to logical sentences in a sensible way
  \cite{DBLP:journals/eccc/GarrabrantBCST16}
\item Semiring programming \cite{DBLP:journals/corr/BelleR16}
\item WMI \cite{DBLP:conf/ijcai/BellePB15}
\item Measures on Boolean algebras
  \begin{itemize}
  \item On possibility and probability measures in finite Boolean algebras
    \cite{DBLP:journals/soco/CastineiraCT02}
  \item Representation of conditional probability measures
    \cite{krauss1968representation}
  \end{itemize}
\end{itemize}

@@

\todo[inline]{This should be replaced with inner sums (a.k.a. free products)}
\begin{definition} \label{def:wmc}
  Let $\mathbf{B}$ be an atomic BA, and let $M \subset \mathbf{B}$ be its set of
  atoms. Let $L \subset \mathbf{B}$ be such that every atom $m \in M$ can be
  uniquely expressed as $m = \bigwedge L'$ for some $L' \subseteq L$, and let
  $w\colon L \to \mathbb{R}_{\ge 0}$ be arbitrary. The \emph{weighted model
    count} $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ is defined as
  \[
    \WMC_w(x) = \begin{cases}
      0 & \text{if } x = 0 \\
      \prod_{l \in L'} w(l) & \text{if } M \ni x = \bigwedge L' \\
      \sum_{\text{atoms } a \le x} \WMC_w(a) & \text{otherwise}
    \end{cases}
  \]
  for any $x \in \mathbf{B}$. Furthermore, we define the \emph{normalised
    weighted model count} $\nWMC_w\colon \mathbf{B} \to [0, 1]$ as $\nWMC_w(x) =
  \frac{\WMC_w(x)}{\WMC_w(1)}$ for all $x \in \mathbf{B}$. For both $\WMC_w$ and
  $\nWMC_w$, we will drop the subscript when doing so results in no potential
  confusion.
%  Finally, we say that a measure $m\colon \mathbf{B} \to
%  \mathbb{R}_{\ge 0}$ is a \emph{WMC measure} (or is \emph{WMC-computable}) if
%  there exists a subset $L \subset \mathbf{B}$ and a weight function $w\colon L
%  \to \mathbb{R}_{\ge 0}$ such that $m = \WMC_w$.
\end{definition}
% TODO: mention that the definition can be reduced to a single formula (i.e.,
% without cases)
% TODO: any measure is a WMC measure if all atoms are in L

@@

\todo[inline]{Outdated. $m(a \land \neg p)$ is better than $m (a \lor p)$.}
\begin{proposition}[Measures on quotients] \label{def:measures_on_quotients}
  Let $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ be a measure, and let $(p)$ be
  a principal ideal. Let $m^*\colon \mathbf{B}/(p) \to \mathbb{R}_{\ge 0}$ be
  defined as
  \[
    m^*(a/(p)) = m(a \lor p)
  \]
  for any $a \in \mathbf{B}$. The function $m^*$ is well-defined. Furthermore,
  it is a measure on $\mathbf{B}/(p)$ if and only if $m(p) = 0$. Moreover, if it
  is a measure, then the following properties transfer from $m$ to $m^*$:
  \begin{itemize}
  \item if $m$ is a \emph{probability measure}, then so is $m^*$;
  \item if $m$ is \emph{strictly positive}, then so is $m^*$.
  \end{itemize}
\end{proposition}
\begin{proof}
  \Cref{lemma:well-defined} proves that the function is well-defined. Next,
  note that
  \[
    m^*(0/(p)) = m(0 \lor p) = m(p),
  \]
  so $m^*(0/(p)) = 0$ if and only if $m(p) = 0$. For the second part of
  \cref{def:measure}, let $a/(p), b/(p) \in \mathbf{B}/(p)$ be such that
  \[
    a/(p) \land b/(p) = a \land b / (p) = 0 / (p).
  \]
  This condition is equivalent to $a \land b \in (p)$ and $(a \land b) \lor p =
  p$ by well-known properties of quotients and ideals
  \cite{givant2008introduction}, \cref{def:ideal}, and the definition of $\le$,
  respectively. Now
  \begin{align*}
    m^*(a/(p) \lor b/(p)) &= m^*(a \lor b / (p)) = m(a \lor b \lor p) = m((a \lor p) \lor (b \lor p)) \\
                          &= m(a \lor p) + m(b \lor p) - m((a \lor p) \land (b \lor p)) \\
                          &= m^*(a/(p)) + m^*(b/(p)) - m((a \lor p) \land (b \lor p))
  \end{align*}
  by \cref{lemma:sum}. However
  \[
    (a \lor p) \land (b \lor p) = (a \land b) \lor p = p,
  \]
  so $m^*(a/(p) \lor b/(p)) = m^*(a/(p)) + m^*(b/(p))$ if and only if $m(p) = 0$.

@@

\todo[inline]{This section made me realise that I was using the wrong definition}

@@

% TODO: start here
%\[
%  \begin{tikzcd}
%    \mathbf{B}/(p) \arrow[rr, shift left, "x/(p) \mapsto x \lor p"] \arrow[rr,
%    shift right, swap, "x/(p) \mapsto x \land \neg p"] & & \mathbf{B} \arrow{r}{\WMC} & \mathbb{R}_{\ge 0}
%  \end{tikzcd}
%\]

%\begin{theorem}[Using the previous results]
%  Let $L \subset \mathbf{B}$ and $w\colon L \to \mathbb{R}_{\ge 0}$ be such that
%  $\WMC_w\colon \mathbf{B} \to \mathbb{R}_{\ge0}$ is a WMC measure. Let $p \in
%  \mathbf{B}$ be such that $\WMC(p) = 0$, and let $g\colon \mathbf{B}/(p) \to
%  \mathbf{B}$ be defined as $g(a/(p)) = a \land \neg p$. Then $\WMC \circ g$
%  is a WMC measure on $\mathbf{B}/(p)$ with respect to the same $L$ and $w$
%  such that
%  \[
%    \begin{tikzcd}
%      \mathbf{B} \arrow{d}[swap]{\WMC} & \mathbf{B}/(p) \arrow{l}[swap]{g}
%      \arrow{dl}{\WMC \circ g} \\
%      \mathbb{R}_{\ge 0}
%    \end{tikzcd}
%  \]
%  commutes.
%\end{theorem}

% TODO: how can WMC compute something in the quotient algebra? Add an example.
% TODO: we can then move through the quotient BA using representatives we get
% from mapping true things to 1 and false things to 0.

@@

\todo[inline]{Proofs need to be updated and propositions could be phrased in a
  better way, but the gist should be the same.}

@@

\section{Polyadic Measure Algebras}

Potential directions to explore:
\begin{itemize}
\item Representing independence and exchangeability. This seems important.
\item (More detail below.) Inequalities as bounds for probabilities. This seems
  to be somewhat explored with other setups.
\item Implementation in SageMath. I would need to define prenex normal forms,
  equality, and lots of other things.
\item Alternative compact ways to define a probability distribution over $N$
  models without assuming that everything is independent. Declaring a measure on
  a FO formula defines a linear equation over the probabilities of models, so
  using this method by itself would require $N-1$ equations, but maybe combining
  this with information about independence and exchangeability can help.
\end{itemize}

\todo[inline]{I'm not sure if it's wise to develop this idea further}

We show how the measure on the models of a FO theory can be extended to a
measure over FO formulas. Two outcomes:
\begin{itemize}
\item For any $p, q \in \mathbf{A}$, if $p \land q = 0$, then $m^*(p \lor q) =
  m^*(p) + m^*(q)$.
\item If $p \le q$, then $m^*(p) \le m^*(q)$. This can be useful in two
  situations:
  \begin{itemize}
  \item If $m^*(q) = 0$ or $m^*(p) = 1$, then this immediately tells us the
    measure on the other sentence.
  \item If we want to find $m^*(p)$, finding $q, r \in \mathbf{A}$ such that $r
    \le p \le q$ bounds the answer.
  \end{itemize}
\end{itemize}

\begin{definition}[\cite{halmos2016algebraic}]
  Given two polyadic algebras $\mathbf{A}$ and $\mathbf{B}$, a \emph{polyadic
    homomorphism} from $\mathbf{A}$ to $\mathbf{B}$ is a Boolean homomorphism
  $f\colon \mathbf{A} \to \mathbf{B}$ such that
  \begin{itemize}
  \item $f\mathbf{S}(\tau)p = \mathbf{S}(\tau)fp$,
  \item $f\bm\exists(J)p = \bm\exists(J)fp$
  \end{itemize}
  for all $\tau \in T$, $p \in \mathbf{A}$, and $J \subseteq I$.
\end{definition}

% by considering several X's, we can have several domains
% I needs to be infinite in order to support arbitrarily complex sentences

\subsection{The Set-Up}

% TODO: Feedback: My general remark here is that this section looks very [???]
% for SRL/WMC/PRM readers. The first-order [???] may not be justified if we are
% only doing essentially propositional models. If we also consider full FoL,
% i.e., infinite domains, and make this a separate paper for KR-type audience.
% That is, let us do the WMC/WMI bit separately and do FO-version in another
% paper for KR track. Title for second paper: "Measures on first-order
% structures via WMC" (or similar...).

\subsubsection{Preliminaries}

What follows is a summary of \cite{halmos2016algebraic}.

% Generic stuff
Let $\mathbf{B}$ be a Boolean algebra (of propositions). Let $X$ be the
(non-empty) domain of discourse. Let $I$ be an index set, elements of which can
be interpreted as variables. The elements of $X^I$ are functions from $I$ to
$X$. For any $x \in X^I$ and $i \in I$, we write $x_i$ to represent $x(i) \in
X$. Let $\mathbf{A^*}$ be the set of all functions $X^I \to \mathbf{B}$, and
note that it forms a Boolean algebra with operations defined pointwise.

% Defining S
Let $T$ be the semigroup of all $I \to I$ transformations. For any $\tau \in
T$, let $\tau_* : X^I \to X^I$ be defined by
\[
  (\tau_* x)_i = x_{\tau i}
\]
for all $x \in X^I$ and $i \in I$. For any (Boolean/polyadic) algebra
$\mathbf{C}$, let $\End(\mathbf{C})$ denote the set of all its endomorphisms. We
can then define $\mathbf{S}$ to be a map $\mathbf{S} : T \to \End(\mathbf{A^*})$
defined by
\[
  \mathbf{S}(\tau)p(x) = p(\tau_* x)
\]
for all $x \in X^I$ and $p \in \mathbf{A^*}$.

% Defining E
For any $J \subseteq I$, let $J_*$ be the relation on $X^I$ defined by
\[
  xJ_*y \quad \iff \quad x_i = y_i \quad \text{for all } i \in I \setminus J
\]
for all $x, y \in X^I$. For any $J \subseteq I$, we then define $\bm\exists(J)$
to be a transformation $\mathbf{A^*} \to \mathbf{A^*}$ defined by
\[
  \bm\exists(J)p(x) = \bigvee_{\substack{y \in X^I,\\ xJ_*y}} p(y)
\]
for all $p \in \mathbf{A^*}$, provided this supremum exists for all $x \in
X^I$\footnote{The universal quantifier $\bm\forall(J)$ is then defined as
  $\bm\forall(J)p = \neg(\bm\exists(J)\neg p)$ for all $p \in \mathbf{A^*}$.}.

Finally, a \emph{functional polyadic (Boolean) algebra}\footnote{To be more
  explicit, a $\mathbf{B}$-valued functional polyadic algebra with domain $X$
  and variables $I$.} is a subalgebra
$\mathbf{A}$ of $\mathbf{A^*}$ such that:
\begin{itemize}
\item $\mathbf{S}(\tau)p \in \mathbf{A}$ for all $p \in \mathbf{A}$ and $\tau
  \in T$;
\item $\bm\exists(J)p \in \mathbf{A}$ for all $p \in \mathbf{A}$ and $J
  \subseteq I$.
\end{itemize}

\begin{proposition}
  For all finite $J, K \subseteq I$, finite $\sigma, \tau \in T$, and all $p \in
  \mathbf{A}$,
  \begin{itemize}
  \item $\bm\exists(\emptyset)p = p$;
  \item $\bm\exists(J)\bm\exists(K) = \bm\exists(J \cup K)$;
  \item $\mathbf{S}(\id)p = p$;
  \item $(\sigma\tau)p = \sigma(\tau p)$;
  \item if $\sigma_{|I \setminus J} = \tau_{|I \setminus J}$, then $\sigma
    \bm\exists(J) = \tau \bm\exists(J)$;
  \item if $\tau$ is injective on $\tau^{-1}J$, then $\bm\exists(J)\tau = \tau
    \bm\exists(\tau^{-1}J)$;
  \item for every $p \in \mathbf{A}$, there exists a finite $J \subseteq I$ such
    that $\bm\exists(i)p = p$ whenever $i \not\in J$.
  \end{itemize}
\end{proposition}

\begin{definition}
  Similarly to $\bm\exists$, a \emph{constant} $c$ is a map $c: \mathcal{P}(I)
  \to \End(\mathbf{A})$ such that:
  \begin{itemize}
  \item $c(\emptyset) = \id_{\mathbf{A}}$;
  \item $c(J \cup K) = c(J)c(K)$;
  \item $c(J)\bm\exists(K) = \bm\exists(K)c(J \setminus K)$;
  \item $\bm\exists(J)c(K) = c(K)\bm\exists(J \setminus K)$;
  \item $c(J)\mathbf{S}(\tau) = \mathbf{S}(\tau)c(\tau^{-1}J)$
  \end{itemize}
  for all $J, K \in \mathcal{P}(I)$ and $\tau \in T$. If $J$ is a singleton
  set $\{ i \}$, we will simply write $c(i)$ instead of $c(J)$.
\end{definition}

\subsubsection{New Results}

\begin{proposition} \label{prop:polyadic_measure}
  Let $\mathbf{B}$ be a finite Boolean algebra with a measure $m :
  \mathbf{B} \to [0, 1]$. Let $\mathbf{A}$ be a $\mathbf{B}$-valued functional
  polyadic algebra with domain $X$ and variables $I$. Let $m^* : \mathbf{A} \to
  \mathbb{R}_{\ge 0}$ be defined by
  \[
    m^*(p) = \sum_{\substack{\text{atoms }y \in \mathbf{B} \text{ s.t.}\\ \exists x \in X^I:\, y \le p(x)}} m(y)
  \]
  for all $p \in \mathbf{A}$. Then $m^*$ is a measure on $\mathbf{A}$.
\end{proposition}

\begin{remark}
  While definitions of $m^*$ such as
  \[
    m^*(p) = m \left( \bigvee_{x \in X^I} p(x) \right)
  \]
  might look tempting, they are not additive.
\end{remark}

\begin{proof} % TODO: Update the proof w.r.t. definitions
  First, we can show that $m^*(1) = 1$ by observing that
  \[
    m^*(1) = \sum_{\text{atoms } y \in \mathbf{B}} m(y) = m \left(
      \bigvee_{\text{atoms } y \in \mathbf{B}} y \right) = m(1) = 1,
  \]
  where we use \cref{thm:representation} and express $1 \in \mathbf{B}$ as the
  supremum of all atoms in $\mathbf{B}$ \cite{ganesh2006introduction}. Clearly
  $m^*(p) \ge 0$ for all $p \in \mathbf{A}$, so we can restrict the codomain of
  $m^*$ to $[0, 1]$.

  Next, we want to show that $m^*(p) > 0$ for all $p \in \mathbf{A} \setminus \{
  0 \}$. If $p \ne 0$, then there must be some $x' \in X^I$ such that $p(x') \ne
  0$. But then, since finite Boolean algebras are atomic, there must also be an
  atom $y \in \mathbf{B}$ such that $y \le p(x')$. Therefore, $m^*(p) \ge m(y) >
  0$, finishing this part of the proof.

  Let $p, q \in \mathbf{A}$ be such that $p \land q = 0$. We want to show
  that $m^*(p \lor q) = m^*(p) \lor m^*(q)$. First, note that
  \[
    y \le (p \lor q)(x) = p(x) \lor q(x)
  \]
  if and only if
  \[
    y = (p(x) \lor q(x)) \land y = (p(x) \land y) \lor (q(x) \land y)
  \]
  by \cref{def:ba}. Also note that
  \[
    (p(x) \land y) \land (q(x) \land y) = p(x) \land q(x) \land y = (p \land
    q)(x) \land y = 0 \land y = 0,
  \]
  so
  \[
    m(y) = m((p(x) \land y) \lor (q(x) \land y)) = m(p(x) \land y) + m(q(x) \land y)
  \]
  by \cref{def:measure} which then leads to
  \begin{align*}
    m^*(p \lor q) &= \sum_{\substack{\text{atoms }y \in \mathbf{B} \text{ s.t.}\\ \exists x \in X^I:\, y \le (p \lor q)(x)}} m(y) = \sum_{\substack{\text{atoms }y \in \mathbf{B} \text{ s.t.}\\ \exists x \in X^I:\, y \le (p \lor q)(x)}} m(p(x) \land y) + m(q(x) \land y) \\
                  &= \sum_{\substack{\text{atoms }y \in \mathbf{B} \text{ s.t.}\\ \exists x \in X^I:\, y \le (p \lor q)(x)}} m(p(x) \land y) + \sum_{\substack{\text{atoms }y \in \mathbf{B} \text{ s.t.}\\ \exists x \in X^I:\, y \le (p \lor q)(x)}} m(q(x) \land y).
  \end{align*}
  Since $y$ is an atom,
  \[
    p(x) \land y  = \begin{cases}
      y & \text{if } y \le p(x) \\
      0 & \text{otherwise,}
    \end{cases}
  \]
  so
  \begin{align*}
    m^*(p \lor q) &= \sum_{\substack{\text{atoms }y \in \mathbf{B} \text{ s.t.}\\ \exists x \in X^I:\, y \le (p \lor q)(x) \text{ and } y \le p(x)}} m(p(x) \land y) + \sum_{\substack{\text{atoms }y \in \mathbf{B} \text{s.t.}\\ \exists x \in X^I:\, y \le (p \lor q)(x) \text{ and } y \le q(x)}} m(q(x) \land y) \\
                  &= \sum_{\substack{\text{atoms }y \in \mathbf{B} \text{ s.t.}\\ \exists x \in X^I:\, y \le p(x)}} m(p(x) \land y) + \sum_{\substack{\text{atoms }y \in \mathbf{B} \text{s.t.}\\ \exists x \in X^I:\, y \le q(x)}} m(q(x) \land y) \\
                  &= \sum_{\substack{\text{atoms }y \in \mathbf{B} \text{ s.t.}\\ \exists x \in X^I:\, y \le p(x)}} m(y) + \sum_{\substack{\text{atoms }y \in \mathbf{B} \text{s.t.}\\ \exists x \in X^I:\, y \le q(x)}} m(y) = m^*(p) + m^*(q),
  \end{align*}
  finishing the proof that $m^*$ is a measure.
\end{proof}

\begin{lemma} \label{lemma:simple_measure}
  Given the setup of \cref{prop:polyadic_measure} and $p \in \mathbf{A}$, if
  $p(x) = p(y)$ for all $x, y \in X^I$ (i.e., $p$ has no free variables), then
  \[
    m^*(p) = m(p(x))
  \]
  (for some $x \in X^I$) is an alternative (i.e., equivalent and simpler)
  definition of $m^*$.
\end{lemma}
\begin{proof}
  Fix some $x \in X^I$. Then
  \[
    m(p(x)) = m \left( \bigvee_{\substack{\text{atoms } y \in \mathbf{B} \text{
            s.t.}\\
          y \le p(x)}} y \right) = \sum_{\substack{\text{atoms } y \in \mathbf{B}
        \text{ s.t.}\\
        y \le p(x)}} m(y) =\sum_{\substack{\text{atoms } y \in \mathbf{B}
        \text{ s.t.}\\
        \exists x' \in X^I : y \le p(x')}} m(y) = m^*(p),
  \]
  where we use \cref{thm:representation} for the first step,
  \cref{def:measure} and \cref{lemma:atomic} for the second step, the
  assumptions of \cref{lemma:simple_measure} for the third step, and the
  definition of $m^*$ for the fourth one.
\end{proof}

\begin{proposition}
  The following identities are true for any $p \in \mathbf{A}$, $J \subseteq I$,
  $\tau \in T$, and constant $c : \mathcal{P}(I) \to \End(A)$:
  \begin{itemize}
  \item $c \neg = \neg c$ (i.e., constants commute with negation);
  \item $\mathbf{S}(\tau) \neg = \neg \mathbf{S}(\tau)$;
  \item $\bm\forall(J)p \le c(J)p \le \bm\exists(J)p$;
  \end{itemize}
\end{proposition}
\begin{proof}
  \begin{itemize}
  \item Obvious.
  \item Ditto.
  \item The first part can be derived from $p \le q \implies \neg q \le \neg p$.
    The second part comes from the definition of $\bm\exists$.
  \end{itemize}
\end{proof}

\subsection{How Probabilities Are Computed}

In order to make the example algebras easily describable, our example programs
will have to be tiny. Consider the following ProbLog
\cite{DBLP:conf/ijcai/RaedtKT07} program:
\begin{align*}
  1.0 &\dblcolon \mathsf{p}(a, b).\\
  0.5 &\dblcolon \mathsf{p}(X, X) \coloneq \mathsf{p}(X, Y);\, \mathsf{p}(Y, X).
\end{align*}
Let $L = \{ \mathsf{p}(a, a), \mathsf{p}(a, b), \mathsf{p}(b, a), \mathsf{p}(b,
b) \}$ be the set of all possible ground atoms. Let $\mathbf{B}$ be the
Boolean algebra freely generated by $L$ (see, e.g.,
\cite{givant2008introduction} for more on free Boolean algebras). Then
$\mathbf{B}$ will have sixteen atoms ranging from $\mathsf{p}(a, a) \land
\mathsf{p}(a, b) \land \mathsf{p}(b, a) \land \mathsf{p}(b, b)$ to
$\neg\mathsf{p}(a, a) \land \neg\mathsf{p}(a, b) \land \neg\mathsf{p}(b, a)
\land \neg\mathsf{p}(b, b)$. The weight function $w : L \to \mathbb{R}_{\ge 0}$
defined by
\[
  w(l) = \begin{cases}
    1 & \text{if } l = \mathsf{p}(a, b) \\
    0.5 & \text{if } l \in \{ \mathsf{p}(a, a), \mathsf{p}(b, b) \} \\
    0 & \text{if } l = \mathsf{p}(b, a) \\
    1-w(l') & \text{if } l = \neg l'
  \end{cases}
\]
for all $l \in L$ defines a WMC measure over $\mathbf{B}$. Note that while we
could define an ideal generated by $\{ \mathsf{p}(b, a), \neg\mathsf{p}(a, b)
\}$ and take the quotient of $\mathbf{B}$ by that ideal to get a Boolean algebra
with a strictly positive measure, this would put zero-probability queries
outside of our algebras, i.e., we would not be able to ask a question whose
answer is zero.

\begin{table}
  \centering
  \caption{Example elements of $\mathbf{A}$ as maps $X^I \to \mathbf{B}$, with
    $a : \mathcal{P}(I) \to \End(\mathbf{A})$ as one of two possible constants.}
  \label{tbl:examples}
  \begin{tabular}{ll}
    \toprule
    Element of $\mathbf{A}$ & Action on $X^I$ \\
    \midrule
    $p = \mathbf{S}(\id)p = \bm\exists(\emptyset)p = a(\emptyset)p = b(\emptyset)p$ & $(x_1, x_2) \mapsto \mathsf{p}(x_1, x_2)$ \\
    $\bm\exists(1)p$ & $(x_1, x_2) \mapsto \mathsf{p}(a, x_2) \lor \mathsf{p}(b, x_2)$ \\
    $\bm\exists(2)p$ & $(x_1, x_2) \mapsto \mathsf{p}(x_1, a) \lor \mathsf{p}(x_1, b)$ \\
    $\bm\exists(I)p$ & $(x_1, x_2) \mapsto \mathsf{p}(a, a) \lor \mathsf{p}(a, b) \lor \mathsf{p}(b, a) \lor \mathsf{p}(b, b)$ \\
    $\mathbf{S}(\{ 1 \mapsto 1, 2 \mapsto 1 \})p$ & $(x_1, x_2) \mapsto \mathsf{p}(x_1, x_1)$ \\
    $\mathbf{S}(\{ 1 \mapsto 2, 2 \mapsto 1 \})p$ & $(x_1, x_2) \mapsto \mathsf{p}(x_2, x_1)$ \\
    $\mathbf{S}(\{ 1 \mapsto 2, 2 \mapsto 2 \})p$ & $(x_1, x_2) \mapsto \mathsf{p}(x_2, x_2)$ \\
    $a(1)p$ & $(x_1, x_2) \mapsto \mathsf{p}(a, x_2)$ \\
    $a(2)p$ & $(x_1, x_2) \mapsto \mathsf{p}(x_1, a)$ \\
    $a(I)p$ & $(x_1, x_2) \mapsto \mathsf{p}(a, a)$ \\
    \bottomrule
  \end{tabular}
\end{table}

Finally, let $\mathbf{A}$ be the functional polyadic algebra $X^I \to
\mathbf{B}$ for $I = \{1, 2\}$ and $X = \{ a, b \}$\footnote{$X$ cannot (or
  should not) have constants that do not occur in $\mathbf{B}$.}. The elements
of $X^I$ can then be represented as tuples $(x_1, x_2)$ for some $x_1, x_2 \in
X$. See \cref{tbl:examples} for example elements of $\mathbf{A}$ which consists
of a single predicate function $p$ and operators $\bm\exists, \mathbf{S}, a, b,
\neg, \land, \lor$, the last three of which are defined pointwise.

\begin{table}
  \centering
  \caption{Step-by-step derivation of how a more complex element of
    $\mathbf{A}$ acts on elements of $X^I$}
  \label{tbl:derivation}
  \begin{tabular}{ll}
    \toprule
    Element of $\mathbf{A}$ & Action on $X^I$ \\
    \midrule
    $p$ & $(x_1, x_2) \mapsto \mathsf{p}(x_1, x_2)$ \\
    $b(2)p$ & $(x_1, x_2) \mapsto \mathsf{p}(x_1, b)$ \\
    $\neg b(2)p$ & $(x_1, x_2) \mapsto \neg\mathsf{p}(x_1, b)$ \\
    $\bm\exists(1)\neg b(2)p$ & $(x_1, x_2) \mapsto \neg\mathsf{p}(a, b) \lor \neg\mathsf{p}(b, b) = \neg(\mathsf{p}(a, b) \land \mathsf{p}(b, b))$ \\
    $\bm\forall(1)b(2)p = \neg\bm\exists(1)\neg b(2)p$ & $(x_1, x_2) \mapsto \neg\neg(\mathsf{p}(a, b) \land \mathsf{p}(b, b)) = \mathsf{p}(a, b) \land \mathsf{p}(b, b)$ \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}
  \centering
  \caption{Atoms $y \in \mathbf{B}$ (and their measures) such that $y \le
    \mathsf{p}(a, b) \land \mathsf{p}(b, b)$}
  \label{tbl:atoms}
  \begin{tabular}{lc}
    \toprule
    Atom $y \in \mathbf{B}$ & $m(y)$ \\
    \midrule
    $\mathsf{p}(a, b) \land \mathsf{p}(b, b) \land \mathsf{p}(a, a) \land \mathsf{p}(b, a)$ & $1 \times 0.5 \times 0.5 \times 0 = 0$ \\
    $\mathsf{p}(a, b) \land \mathsf{p}(b, b) \land \neg\mathsf{p}(a, a) \land \mathsf{p}(b, a)$ & $1 \times 0.5 \times 0.5 \times 0 = 0$ \\
    $\mathsf{p}(a, b) \land \mathsf{p}(b, b) \land \mathsf{p}(a, a) \land \neg\mathsf{p}(b, a)$ & $1 \times 0.5 \times 0.5 \times 1 = 0.25$ \\
    $\mathsf{p}(a, b) \land \mathsf{p}(b, b) \land \neg\mathsf{p}(a, a) \land \neg\mathsf{p}(b, a)$ & $1 \times 0.5 \times 0.5 \times 1 = 0.25$ \\
    \bottomrule
  \end{tabular}
\end{table}
